Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-08-17 surfaces-write-up/surface-write-up.tex
edits to my prevoius edits

\maketitle
\todo{Aaron: Remove labels which are not referenced}
\todo{Aaron: Make sure indexing at 0 is correct in projective space
section, particularly toward the end}
%%%%%%%%%%%%%%%%%%%%%%%%%%%% Introduction %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\noindent
In Section \ref{sec:hirz}, we shift our attention to Hirzebruch Surfaces. Recall that for each $m \geq 0$ we can define the
Hirzebruch surface $F_m := \proj(\sym^\bullet \sco_{\bp^1} \oplus
\sco_{\bp^1}(m)),$ or as a quotient of an open subscheme of
$\ba^4$ by a 
kweighted action, as is done in the
beginning of Section \ref{sec:hirz}.
Note that Hirzebruch surfaces are normal
and locally factorial, being $\bp^1$ bundles over
$\bp^1$, and therefore the natural map from Cartier
divisors to Weil divisors is an isomorphism.
In other words, all Cartier divisors arise as the zero locus
of a section of some line bundle. The analogous fact for
projective space was also implicitly used in Theorems
~\ref{thm:proj-effective-intro} and ~\ref{thm:proj-generators-relations}.

@@

\ssec{Notations and Conventions}
Let $D$ be a (Weil) $\bq$-divisor of the form
\begin{align*}
	D = \sum_{i=1}^{n}\alpha_i D_i \in \di X \otimes \bq.
\end{align*}

\noindent
where $\alpha_i \in \bq$ and $D_i \in \di X$ is an integral
codimension 1 closed subscheme of $X$. When it is convenient to do
so, we shall sometimes start the indexing at $0$, so that $i$ runs
from $0$ to $n$. Define the degree of $D$ by $\deg D \colon =
\sum_{i=1}^{n}\alpha_i \cdot \deg D_i$. For such a divisor $D$ with $\alpha_i =
\frac{c_i}{k_i}$ let $\ell_i = \lcm_{j\ne i}(k_j)$
and let $\ell_{i,j} = \lcm_{h\ne i,j}(k_h).$ The floor of a $\bq$-divisor $D$ is a divisor $\lfloor D
\rfloor \in \di X$ is defined to be $\lfloor D \rfloor \colon = \sum_{i = 1}^{n}
\lfloor \alpha_i \rfloor P_i$. Let $
	R_D \colon = \bigoplus_{d \geq 0} u^d H^0(\bp^m, \lfloor dD \rfloor)$
denote the canonical ring associated to the $\bq$-divisor $D$. If $R$
is a graded ring, we denote the $d$th graded component of $R$ by $R_d$. If $r$ is a rational number, we let
$\fr(r) \colon = r - \lfloor r
\rfloor$ denote the fractional part of $r$. Throughout, we
shall work over a fixed base field $\bk$.

@@

For the remainder of this section, we shall make the additional assumption that 
\begin{align} \label{eqn:pm-basis-assumption} f_0, \ldots, f_{m}
	\text{ are $m$ independent linear forms.} \end{align}

>>>>

\maketitle
\todo{Aaron: Remove labels which are not referenced}
\todo{Aaron: Make sure indexing at 0 is correct in projective space
section, particularly toward the end}
\todo{Aaron: Make sure to use $\bk$ instead of $k$ for the field}
%%%%%%%%%%%%%%%%%%%%%%%%%%%% Introduction %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\noindent
In Section \ref{sec:hirz}, we shift our attention to Hirzebruch surfaces. Recall that for each $m \geq 0$ we can define the
Hirzebruch surface $F_m := \proj(\sym^\bullet \sco_{\bp^1} \oplus
\sco_{\bp^1}(m)),$ or as a quotient of an open subscheme of
$\ba^4$ by a 
weighted action, as is done in the
beginning of Section \ref{sec:hirz}.
Note that Hirzebruch surfaces are normal
and locally factorial, being $\bp^1$ bundles over
$\bp^1$, and therefore the natural map from Cartier
divisors to Weil divisors is an isomorphism.
In other words, all Cartier divisors arise as the zero locus
of a section of some line bundle. We implicitly use this in the
statement of the following theorem. The analogous fact for
projective space was also implicitly used in Theorems
~\ref{thm:proj-effective-intro} and ~\ref{thm:proj-generators-relations}.

@@

\ssec{Notations and Conventions}
We will use $m \in \bz$ to index the dimension of a given projective
space $\bp^m$ or the type of the hirzebruch surface $\hirz_m$.
Let $D$ be a (Weil) $\bq$-divisor of the form
\begin{align*}
	D = \sum_{i=1}^{n}\alpha_i D_i \in \di X \otimes \bq.
\end{align*}

\noindent
where $n \in \bz$ indexes the number of irreducible divisors in 
the above expansion of $D$, $\alpha_i \in \bq$, and $D_i \in \di X$ is an integral
codimension 1 closed subscheme of $X$. When it is convenient to do
so, we shall sometimes start the indexing at $0$, so that $i$ runs
from $0$ to $n$. Define the degree of $D$ by $\deg D \colon =
\sum_{i=1}^{n}\alpha_i \cdot \deg D_i$. For such a divisor $D$ with $\alpha_i =
\frac{c_i}{k_i}$ let $\ell_i = \lcm_{j\ne i}(k_j)$
and let $\ell_{i,j} = \lcm_{h\ne i,j}(k_h).$ The floor of a $\bq$-divisor $D$ is a divisor $\lfloor D
\rfloor \in \di X$ is defined to be $\lfloor D \rfloor \colon = \sum_{i = 1}^{n}
\lfloor \alpha_i \rfloor P_i$. Let $
	R_D \colon = \bigoplus_{d \geq 0} u^d H^0(\bp^m, \lfloor dD \rfloor)$
denote the canonical ring associated to the $\bq$-divisor $D$. If $R$
is a graded ring, we denote the $d$th graded component of $R$ by $R_d$. If $r$ is a rational number, we let
$\fr(r) \colon = r - \lfloor r
\rfloor$ denote the fractional part of $r$. Throughout, we
shall work over a fixed base field $\bk$.

@@

For the remainder of this section, we shall make the additional assumption that 
\begin{align} \label{eqn:pm-basis-assumption} f_0, \ldots, f_{m}
	\text{ are independent linear forms.} \end{align}
